Early range shifts in transboundary fish stocks challenge their management under climate change

### 2021-03-16 Scripts/Manuscript.Rmd
Updated manuscript with title and Alessandro's comments

---
title: Identifying timing of shifts in transboundary fish stocks to support climate resilient fisheries management

author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland. ^4^School of Environmental Sciences, University of Liverpool. ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
output:
  bookdown::word_document2
reference_docx:  ../References/manuscript_template.docx
  # pdf_document: default
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
header-includes:
- \usepackage{amsmath}
- \usepackage{float} # Use the 'float' package
- \floatplacement{figure}{H} # Make every figure with caption = h
---

@@

Early range shifts in transboundary fish stocks challenge their management under climate change

@@

Different levels of uncertainty exist in the projected emergence of environmental variables, mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg]. First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species. Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg]. Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution. This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to the emergence of environmental drivers is larger but that to scenarios is lower [@Frolicher:2016dg]. While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq]. As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade. This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, and where internal variability (e.g., ENSO for Pacific nations) accounts for most of the uncertainty. Model uncertainty also affects the way we project species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh]. Exercises looking at changes in fish biomass with multiple models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh]. Further research that includes large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce]. Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks. While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca]. Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie]. Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.

>>>>

---
title: "Early range shifts in transboundary fish stocks challenge their management under climate change"
author: 'Authors: Juliano Palacios-Abrantes^1*^, Thomas Frölicher^2,3^, Gabriel Reygondeau^1^, U. Rashid Sumaila^1^, Alessandro Tagliabue^4^, Colette Wabnitz^1,5^, and William W.L. Cheung^1^'
date: 'Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Climate and Environmental Physics, Physics Institute, University of Bern. ^3^Oeschger Centre for Climate Change Research, University of Bern, Switzerland. ^4^School of Environmental Sciences, University of Liverpool. ^5^Stanford Center for Ocean Solutions, Stanford University, Stanford, United States'
output: bookdown::word_document2
reference_docx:  ../References/sci_template.docx
  # pdf_document: default
bibliography: ../References/ToE_Ref.bib
csl: ../References/science.csl
---

@@

Rapid climate-induced shifts in transboundary fish stocks

@@

Different levels of uncertainty exist in the projected emergence of environmental variables mainly related to *i*) climate change scenarios and  *ii*) model structure [@Frolicher:2016dg]. First, our analysis is based on a high emission scenario (RCP 8.5) representing an extreme case of shifts in the shared distribution of transboundary species. Thus, any climate change mitigation pathway (e.g., RCP 2.6) could result in a substantial delay in distribution shifts as environmental signals are sensitive to mitigation, at least after the middle of the $21^{st}$ century [@Frolicher:2016dg]. Second, a substantial source of uncertainty is related to model selection for both climate change and fisheries distribution. This is specifically important for early time periods (e.g., 2016-2035) where the uncertainty related to the emergence of environmental drivers is larger but that to scenarios is lower [@Frolicher:2016dg]. While such uncertainty could potentially be reduced by incorporating  ensemble simulations from a range of different ESMs, these are computationally expensive simulations [@Frolicher:2009ge; @Frolicher:2016dg; @Rodgers:2015jq] and they ignore other sources of structural uncertainty based on the parameterisation of processes. As a consequence, transboundary fisheries managers are facing changes in stocks’ shared distributions embedded in complex uncertainties, and will likely continue to do so in the coming decade. This is critical for the tropics, where the shared distribution of transboundary stocks is expected to shift the earliest, the response of the base of the food web to climate change is most uncertain [@Tagliabue:2020hd; @Kwiatkowski:2020cf] and where internal variability (e.g., El Niño Southern Oscillation - ENSO- for Pacific nations) is also important . These uncertainties also affect the projection of species distributions [@Lotze:2019ce; @BryndumBuchholz:2019kh]. Overall, changes in fish biomass with multiple upper trophic level models, including the DBEM, show overall agreement across models in terms of direction of change, but variability in the magnitude of that change [@BryndumBuchholz:2019kh]. Further research that reduces uncertainty in the response of the base of ocean food webs, alongside large ensemble simulations of multiple species distribution models could elucidate a broader range in structural uncertainty of fish and fisheries models [@Lotze:2019ce]. Another important source of uncertainty in this study is the utilization of political boundaries (EEZ) to define stocks. While this method might define some stocks that do not necessarily align with biologically-defined sub-populations within an EEZ, in many EEZs, fisheries are often managed at the species level [@MAP:2017uf] and sub-populations are potentially interconnected [@Ramesh:2019va], thus, providing additional ecological ground for our analysis [@Dunn:2019ca]. Reproducing our analysis regionally, where spatially explicit stock data is available would allow to generate better constrained results in changes in both time of emergence and stock share ratio of transboundary stocks, and potentially identity different types of shared stock shifts at meta-population level [@Link:2010ei; @Archambault:2016ie]. Addressing these uncertainties systematically can serve as a roadmap for future studies to provide additional information to inform policy towards sustainable and equitable international fisheries management under climate change.
